Accelerating Optimal Power Flow with GPUs: SIMD Abstraction of Nonlinear Programs and Condensed-Space Interior-Point Methods

### 2023-07-28 paper/main.tex
changes

\date{\small
  $^{1}$Mathematics and Computer Science Division, Argonne National Laboratory\\
  $^{2}$Centre Automatique et Syst\`{e}mes, Mines Paris - PSL
}
\begin{document}
\maketitle
\begin{abstract}
  This paper introduces a novel computational framework for solving
alternating current optimal power flow (ACOPF) problems using
graphics processing units (GPUs). While GPUs have demonstrated
remarkable performance in various computing domains, their application
in AC OPF has been limited due to challenges associated with porting
sparse automatic differentiation (AD) and sparse linear solver
routines to GPUs. We aim to address these issues with two key
strategies. First, we utilize a single-instruction, multiple-data
(SIMD) abstraction of nonlinear programs (NLP). This approach enables
the specification of model equations while preserving their
parallelizable structure, and in turn, facilitates the implementation
of AD routines that can exploit such structure. Second, we employ a
condensed-space interior-point method (IPM) with an inequality
relaxation strategy. This technique involves relaxing equality
constraints to inequalities and condensing the Karush-Kuhn-Tucker
system into a much smaller positive definite matrix. This strategy
offers the key advantage of being able to factorize the KKT matrix
without numerical pivoting, which in the past has hampered the parallelization of
the IPM algorithm. By combining these two strategies, we can perform the
majority of operations on GPUs while keeping the data residing in the
device memory only. Comprehensive numerical benchmark results showcase
the substantial computational advantage of our approach. Remarkably,
for solving large-scale AC OPF problems to a moderate accuracy, our
implementations—MadNLP.jl and SIMDiff.jl—running on NVIDIA GPUs
achieve an order of magnitude speedup compared to state-of-the-art
tools running on contemporary CPUs.
\end{abstract}

@@

\paragraph*{Contributions}
We present, for the first time, a nonlinear optimization framework
that can run entirely on GPU, with all the performance-critical data
arrays residing exclusively on GPUs. Additionally, we introduce the
concept of SIMD abstraction for NLP problems, which results in an
efficient implementation of GPU-accelerated automatic differentiation,
leading to a 50x speedup compared to existing tools for large-scale
OPFs. Furthermore, we propose the condensed IPM with inequality
relaxation strategy for the first time, enabling the treatment of the
KKT systems of equality-constrained NLPs without numerical pivoting,
thus allowing the solution of sparse, large-scale NLPs (with a
prominent example being AC OPFs) on GPUs.

\paragraph*{Related Work}
Several recent works have explored the use of GPUs for large-scale
nonlinear optimization problems. Cao et al. \cite{cao2016augmented}
proposed an augmented Lagrangian interior-point approach that employs
augmented Lagrangian outer iteration and the treatment of linear
systems using a preconditioned conjugate gradient method.  Another
recent development is the hybrid (direct-iterative) KKT system solver
specifically designed for GPUs \cite{regev2023hykkt}.  Before the
introduction of the condensed space IPM with an inequality relaxation
strategy, the authors have investigated the use of reduction
strategies (state variable elimination) to treat KKT systems in a
dense form on the GPU
\cite{pacaud2023parallel,pacaud2022feasible,pacaud2023accelerating,cole2023exploiting}.
In parallel, approaches based on Lagrangian decomposition and
batched nonlinear optimzition solver based on TRON algorithm \cite{lin1999newton}
has been investigated \cite{kim2022accelerated,kim2021leveraging}.
An NLP solver for high-performance computers (HPC) with GPU
accelerators called HiOP has been under development
\cite{hiop_techrep}, with a similar scope as our solver MadNLP.jl.
The implementation of derivative evaluations with the exploitation of
repeated structures within model equations has, to the best of our
knowledge, been first introduced in Gravity \cite{Gravity}. This was
achieved through the introduction of so-called template constraints,
and multi-threaded derivate evaluation has been implemented therein.
However, it is important to note that their differentiation approach
is based on symbolic differentiation.

\paragraph*{Organization}
The paper is organized as follows. In the remainder of the current
section, we introduce the mathematical notation. In Section
\ref{sec:prelim}, we provide general preliminary knowledge on
numerical optimization and GPU computing. In Section \ref{sec:simd},
we present the SIMD abstraction of NLPs and their advantages in terms
of implementing parallel AD. Section \ref{sec:ipm} presents the
optimization algorithm under study, the condensed-space IPM with an
inequality relaxation strategy. Section \ref{sec:num} presents the
numerical results, comparing our approach with other state-of-the-art
solution methods on the CPUs. Finally, conclusions and future outlooks
are given in Section \ref{sec:conc}.

\paragraph*{Notation}
We denote the set of real numbers and the set of integers by
$\mathbb{R}$ and $\mathbb{I}$. We let $[M]:=\{1,2,\cdots,M\}$. We let
$[v_i]_{i\in[M]}:=[v_1;v_2;\cdots,v_M]$.  A vector of ones with an
appropriate size is denoted by $\boldsymbol{1}$. An identity matrix
with an appropriate size is denoted by $I$. For matrices $A$ and $B$,
$A\succ(\succeq) B$ indicates that $A-B$ is positive (semi)-definite
while $A>(\geq) B$ denotes a component-wise inequality. We use the
following convention for any symbol $x$: $X:=\diag(x)$.

@@

In contrast, AD differentiates computer
programs directly by inspecting the computation graph and applying chain rules,
to evaluate derivatives efficiently and accurately.
This approach has become the dominant paradigm for derivative computation within the
scientific computing domain, including NLP and
machine learning. For large-scale optimization problems, such as AC OPFs, AD
tools are often implemented as part of domain-specific modeling
languages, such as JuMP, CasADi, and AMPL (optimization) and Tensorflow, Torch and
Flux (machine learning).

@@

The Julia Language, our language of choice, offers convenient and
efficient ways to implement automatic differentiation. Through the use
of the multiple dispatch paradigm~\cite{bezanson2017julia} \textit{any Julia function} ---
including commonly used operations like addition, multiplication,
trigonometric and exponential functions (among others) --- can be easily
overloaded. Multiple dispatch allows functions to be dynamically
dispatched based on the run-time type, a crucial feature for
implementing differentiable programming. Several AD implementations
have been developed in Julia Language, such as ReverseDiff.jl,
ForwardDiff.jl, Zygote.jl, and JuMP.jl. While these tools are general
and useful for various applications, they are not optimized for
evaluating derivatives of AC OPF problems, as they are not designed to
exploit the parallelizable structures in the model.

\subsection{GPU Computing}\label{sec:gpu}
With the increasing prevalence of GPU in various scientific computing
domains, there has been growing interest in leveraging these emerging
architectures to efficiently solve large-scale NLPs, like AC OPF problems.
However, adapting a NLP solution algorithms
such as IPM to GPUs presents challenges due to the fundamental
differences between GPU and CPU programming paradigms. While CPUs
execute a sequence of instructions on a single input (single
instruction, single data, or SISD, in Flynn's taxonomy), GPUs run the
same instruction simultaneously on hundreds of threads using the SIMD
paradigm (see Fig. \ref{fig:simd}). The SIMD parallelism works well
for algorithms that can be decomposed into simple instructions running
entirely in parallel, but not all algorithms fit this
paradigm. For example, branching in the control flow can hinder
lockstep execution across multiple threads, and in turn, prevent
efficient implementations on GPUs. On the contrary, when the
algorithm's structure allows for efficient parallelization, the SIMD
parallelism in GPUs can offer orders of magnitudes speedup.

We highlight that the following, arguably common, computational
patterns are particularly effective when implemented on GPUs:
\begin{align}
  y&\leftarrow \left[\phi(x; q_j)\right]_{j\in [J]}\tag{Pattern 1}\label{eqn:pattern-1}\\
  o&\leftarrow  \mathop{\rm{Op}}_{i\in [I]} \psi(x;p_i) \tag{Pattern 2}\label{eqn:pattern-2}\\
  x&\leftarrow  \chi_{s_1}\circ\cdots\circ \chi_{s_K}(x) \tag{Pattern 3}\label{eqn:pattern-3}
\end{align}
Here, $\psi:\mathbb{R}^{n_x}\times \mathbb{R}^{n_{p}}\rightarrow
\mathbb{R}$ , $\phi:\mathbb{R}^{n_x}\times
\mathbb{R}^{n_{q}}\rightarrow \mathbb{R}$, and
$\chi:\mathbb{R}^{n_x}\times \mathbb{R}^{n_{s}}\rightarrow
\mathbb{R}^{n_x}$ are simple instructions that require only a small number
of operations; $\mathop{\rm{Op}}$ is a monoid operator on
$\mathbb{R}\cup\{+\infty,-\infty\}$, such as addition, multiplication,
maximum, and minimum. In \ref{eqn:pattern-3}, we denote $\chi_{s_k}(x):=\chi(x,s_k)$ and
assume that $\circ$ is commutative for $\{\chi_{s_k}(\cdot)\}_{\forall
s_k}$. \ref{eqn:pattern-1} is typically most effective on GPUs, where
each thread employed can operate independently without needing to
simultaneously manipulate the same device memory location. While
\ref{eqn:pattern-2} and \ref{eqn:pattern-3} are less effective,
they still can be significantly
faster than operations in CPUs, as substantial part of the operation
can still be parallelized by the use of buffers. In simple cases, the
implementation of these operations can be performed with the standard
{\tt map} and {\tt mapreduce} programming models. However, in more
complex cases, especially for \ref{eqn:pattern-3}, the
implementation may require pre-inspection of memory write-access
patterns and the use of custom kernels.

Many of the operations required in AD of sparse physical models, as
well as the application of optimization algorithms, are based on the
computational patterns mentioned above. For example, an AC OPF model can
be implemented with 15 different computational patterns (see Section~\ref{sec:simd:ad}), most of which
fall within the aforementioned categories. Furthermore, the
computation within optimization solvers, such as forming the
left-hand-side for the KKT systems, computing the $\|\cdot\|_\infty$
norm of the constraint violation, and assembling the condensed KKT
system, can be carried out by using these computational patterns as well. The
only exception is the factorization of the sparse KKT matrix, which
requires more sophisticated implementations.

@@

\subsection{Abstraction}
The SIMD abstraction under consideration is as follows:
\begin{align}\label{eqn:prob}
  \min_{x^\flat\leq x \leq x^\sharp}
  & \sum_{l\in[L]}\sum_{i\in [I_l]} f^{(l)}(x; p^{(l)}_i)\\\nonumber
  \st\; &\forall m\in[M]:\\\nonumber
  &\begin{aligned}[t]
    \left[g^{(m)}(x; q_j)\right]_{j\in [J_m]} +\sum_{n\in [N_m]}\sum_{k\in [K_n]}h^{(n)}(x; s^{(n)}_{k}) =0,
  \end{aligned}
  \end{align}
where $f^{(\ell)}(\cdot,\cdot)$, $g^{(m)}(\cdot,\cdot)$, and
$h^{(n)}(\cdot,\cdot)$ are twice differentiable functions with respect
to the first argument, whereas $\{\{p^{(k)}_i\}_{i\in [N_k]}\}_{k\in[K]}$,
$\{\{q^{(k)}_{i}\}_{i\in [M_l]}\}_{m\in[M]}$, and
$\{\{\{s^{(n)}_{k}\}_{k\in[K_n]}\}_{n\in[N_m]}\}_{m\in[M]}$ are
problem data, which can either be discrete or continuous.  We assume
that our functions $f^{(l)}(\cdot,\cdot)$, $g^{(m)}(\cdot,\cdot)$, and
$h^{(n)}(\cdot,\cdot)$ can be expressed with computational
graphs of moderate length. One can observe that the problem in
\eqref{eqn:prob} is expressed by the computational patterns in Section
\ref{sec:gpu}. In particular, the objective function falls within \ref{eqn:pattern-1},
the first term in the constraint falls within \ref{eqn:pattern-2}, and the second term in the constraint falls within \ref{eqn:pattern-3}.
Accordingly, the evaluation and differentiation of the
model equations in \eqref{eqn:prob} is amenable to SIMD parallelism.

@@

\subsection{Parallel AD}
\label{sec:simd:ad}
Many physics-based models, such as AC OPF, have a highly repetitive
structure. One of the manifestations of it is that the mathematical
statement of the model is concise, even if the practical model may contain
millions of variables and constraints. This is possible due to the use of
repetition over a certain index and data sets. For example,
it suffices to use 15 computational patterns to fully specify the
AC OPF model. These patterns arise from (1) generation cost, (2) reference
bus voltage angle constraint, (3-6) active and reactive power flow (from and to),
(7) voltage angle difference constraint, (8-9) apparent
power flow limits (from and to), (10-11) power balance equations,
(12-13) generators' contributions to the power balance equations, and
(14-15) in/out flows contributions to the power balance
equations. However, such repetitive structure is not well exploited
in the standard NLP modeling paradigms. In
fact, without the SIMD abstraction, it is difficult for the AD package
to detect the parallelizable structure within the model, as it will
require the full inspection of the computational graph. By preserving
the repetitive structures in the model, one can facilitate the
implementation of parallelized derivative computations.

>>>>

\date{\small
  $^{1}$Mathematics and Computer Science Division, Argonne National Laboratory\\
  $^{2}$Centre Automatique et Syst\`{e}mes, Mines Paris - PSL
}
\begin{document}
\maketitle
\begin{abstract}
  This paper introduces a novel computational framework for solving
alternating current optimal power flow (ACOPF) problems using
graphics processing units (GPUs). While GPUs have demonstrated
remarkable performance in various computing domains, their application
in AC OPF has been limited due to challenges associated with porting
sparse automatic differentiation (AD) and sparse linear solver
routines to GPUs. We aim to address these issues with two key
strategies. First, we utilize a single-instruction, multiple-data
(SIMD) abstraction of nonlinear programs (NLP). This approach enables
the specification of model equations while preserving their
parallelizable structure, and in turn, facilitates the implementation
of AD routines that can exploit such structure. Second, we employ a
condensed-space interior-point method (IPM) with an inequality
relaxation strategy. This technique involves relaxing equality
constraints to inequalities and condensing the Karush-Kuhn-Tucker
system into a much smaller positive definite system. This strategy
offers the key advantage of being able to factorize the KKT matrix
without numerical pivoting, which in the past has hampered the parallelization of
the IPM algorithm. By combining these two strategies, we can perform the
majority of operations on GPUs while keeping the data residing in the
device memory only. Comprehensive numerical benchmark results showcase
the substantial computational advantage of our approach. Remarkably,
for solving large-scale AC OPF problems to a moderate accuracy, our
implementations—MadNLP.jl and SIMDiff.jl—running on NVIDIA GPUs
achieve an order of magnitude speedup compared to state-of-the-art
tools running on contemporary CPUs.
\end{abstract}

@@

\paragraph*{Contributions}
We present, for the first time, a nonlinear optimization framework
that can run entirely on GPU, with all the performance-critical data
arrays residing exclusively on GPUs. Additionally, we introduce the
concept of SIMD abstraction for NLP problems, which results in an
efficient implementation of GPU-accelerated parallel AD, leading to a
50x speedup compared to existing tools for large-scale
OPFs. Furthermore, we propose the condensed IPM with inequality
relaxation strategy for the first time, enabling the treatment of the
KKT systems of NLPs without numerical pivoting, thus allowing the
solution of sparse, large-scale NLPs (with a prominent example being
AC OPFs) on GPUs.

\paragraph*{Related Work}
Several recent works have explored the use of GPUs for large-scale
nonlinear optimization problems. Cao et al. \cite{cao2016augmented}
proposed an augmented Lagrangian interior-point approach that employs
augmented Lagrangian outer iteration and the treatment of linear
systems using a preconditioned conjugate gradient method. Prior to the
introduction of the sparse condensed-space IPM with an inequality relaxation
strategy, the authors have investigated the use of reduction
strategies (state variable elimination) to treat KKT systems in a
dense form on the GPU
\cite{pacaud2023parallel,pacaud2022feasible,pacaud2023accelerating}.
In parallel, approaches based on Lagrangian decomposition and
batched with batched TRON solver \cite{lin1999newton}
has been investigated \cite{kim2022accelerated,kim2021leveraging}.
An NLP solver for high-performance computers (HPC) with GPU
accelerators called HiOP has been under development
\cite{hiop_techrep}, with a similar scope as our solver MadNLP.jl.
 Another recent development is the hybrid (direct-iterative) KKT
system solver specifically designed for GPUs \cite{regev2023hykkt}.
The implementation of derivative evaluations with the exploitation of
repeated structures within model equations has, to the best of our
knowledge, been first introduced in Gravity \cite{Gravity}. This was
achieved through the introduction of so-called template constraints,
and multi-threaded derivate evaluation has been implemented therein.
However, it is important to note that their differentiation approach
is based on symbolic differentiation.
The idea of condensed-space IPM (without inequlaity relaxation
strategy) is not new, but they have been used primarily in more
specific contexts, where the increased nonzero entries in the KKT system
is less of a concern, such as in the dense form model predictive control
problems \cite{jerez2012sparse,cole2023exploiting}.

\paragraph*{Organization}
The paper is organized as follows. In the remainder of the current
section, we introduce the mathematical notation. In Section
\ref{sec:prelim}, we provide general preliminary knowledge on
numerical optimization, automatic differentiation, and GPU computing.
In Section \ref{sec:simd}, we present the SIMD abstraction of NLPs and
their advantages in terms of implementing parallel AD. Section
\ref{sec:ipm} presents the optimization algorithm under study, the
condensed-space IPM with an inequality relaxation strategy. Section
\ref{sec:num} presents the numerical results, comparing our approach
with other state-of-the-art solution methods on the CPUs. Finally,
conclusions and future outlooks are given in Section \ref{sec:conc}.

\paragraph*{Notation}
We denote the set of real numbers and the set of integers by
$\mathbb{R}$ and $\mathbb{I}$. We let $[M]:=\{1,2,\cdots,M\}$. We let
$[v_i]_{i\in[M]}:=[v_1;v_2;\cdots,v_M]$.  A vector of ones with an
appropriate size is denoted by $\boldsymbol{1}$. An identity matrix
with an appropriate size is denoted by $I$. For matrices $A$ and $B$,
$A\succ(\succeq) B$ indicates that $A-B$ is positive (semi)-definite
while $A>(\geq) B$ denotes a component-wise inequality. We use the
convention $X:=\diag(x)$ for any symbol $x$.

@@

In contrast, AD differentiates computer
programs directly by inspecting the computation graph and applying chain rules,
to evaluate derivatives efficiently and accurately.
This approach has become the dominant paradigm for derivative computation within the
scientific computing domain, including NLP and
machine learning. For large-scale optimization problems, such as AC
OPFs, AD tools are often implemented as part of domain-specific
modeling languages. Examples of such modeling languages include JuMP,
CasADi, and AMPL (optimization) and Tensorflow, Torch and, Flux
(machine learning).

@@

The Julia Language, our language of choice, offers convenient and
efficient ways to implement automatic differentiation. Through the use
of the multiple dispatch paradigm~\cite{bezanson2017julia} \textit{any
Julia function}---including commonly used operations like addition,
multiplication, trigonometric and exponential functions (among
others)---can be easily overloaded. Multiple dispatch allows functions
to be dynamically dispatched based on the run-time type, a crucial
feature for implementing differentiable programming. Several AD
implementations have been developed in Julia Language, such as
ReverseDiff.jl, ForwardDiff.jl, Zygote.jl, and JuMP.jl. While these
tools are general and useful for various applications, they are not
optimized for evaluating derivatives of AC OPF problems, as they are
not designed to exploit the parallelizable structures in the model,
while preserving the desired sparsity.

\subsection{GPU Computing}\label{sec:gpu}
With the increasing prevalence of GPU in various scientific computing
domains, there has been growing interest in leveraging these emerging
architectures to efficiently solve large-scale NLPs, like AC OPF problems.
However, adapting a NLP solution algorithms,
such as IPM, to GPUs presents challenges due to the fundamental
differences between GPU and CPU programming paradigms. While CPUs
execute a sequence of instructions on a single input (single
instruction, single data, or SISD, in Flynn's taxonomy), GPUs run the
same instruction simultaneously on hundreds of threads using the SIMD
paradigm (see Fig. \ref{fig:simd}). The SIMD parallelism works well
for algorithms that can be decomposed into simple instructions running
entirely in parallel, but not all algorithms fit this
paradigm. For example, branching in the control flow can hinder
lockstep execution across multiple threads, and in turn, prevent
efficient implementations on GPUs. On the contrary, when the
algorithm's structure allows for efficient parallelization, the SIMD
parallelism in GPUs can offer orders of magnitudes speedup.

We highlight that the following, arguably common, computational
patterns are particularly effective when implemented on GPUs:
\begin{align}
  y&\leftarrow \left[\phi(x; q_j)\right]_{j\in [J]}\tag{Pattern 1}\label{eqn:pattern-1}\\
  o&\leftarrow  \mathop{\rm{Op}}_{i\in [I]} \psi(x;p_i) \tag{Pattern 2}\label{eqn:pattern-2}\\
  x&\leftarrow  \chi_{s_1}\circ\cdots\circ \chi_{s_K}(x) \tag{Pattern 3}\label{eqn:pattern-3}
\end{align}
Here, $\psi:\mathbb{R}^{n_x}\times \mathbb{R}^{n_{p}}\rightarrow
\mathbb{R}$ , $\phi:\mathbb{R}^{n_x}\times
\mathbb{R}^{n_{q}}\rightarrow \mathbb{R}$, and
$\chi:\mathbb{R}^{n_x}\times \mathbb{R}^{n_{s}}\rightarrow
\mathbb{R}^{n_x}$ are simple instructions that require only a small number
of operations; $\mathop{\rm{Op}}$ is a monoid operator on
$\mathbb{R}\cup\{+\infty,-\infty\}$, such as addition, multiplication,
maximum, and minimum. In \ref{eqn:pattern-3}, we denote $\chi_{s_k}(x):=\chi(x,s_k)$ and
assume that $\circ$ is commutative for $\{\chi_{s_k}(\cdot)\}_{\forall
s_k}$. \ref{eqn:pattern-1} is typically most effective on GPUs, where
each thread employed can operate independently without needing to
simultaneously manipulate the same device memory location. While
\ref{eqn:pattern-2} and \ref{eqn:pattern-3} are less effective,
they still can be significantly
faster than operations on CPUs, as substantial part of the operation
can still be parallelized by the use of buffers. In simple cases, the
implementation of these operations can be performed with the standard
{\tt map} and {\tt mapreduce} programming models. However, in more
complex cases, especially for \ref{eqn:pattern-3}, the implementation
may require pre-inspection of memory write-access patterns and the use
of custom kernels.

Many of the operations required in AD of sparse physical models, as
well as the application of optimization algorithms, are based on the
computational patterns mentioned above. For example, an AC OPF model
can be implemented with 15 different computational patterns (see
Section~\ref{sec:simd:ad}), all of which fall within the
aforementioned categories. Furthermore, the computation within
optimization solvers, such as forming the left-hand-side for the KKT
systems, computing the $\|\cdot\|_\infty$ norm of the constraint
violation, and assembling the condensed KKT system, can be carried out
by using these computational patterns as well. The only exception is
the factorization of the sparse KKT matrix, which requires more
sophisticated implementations.

@@

\subsection{Abstraction}
The SIMD abstraction under consideration is as follows:
\begin{align}\label{eqn:prob}
  \min_{x^\flat\leq x \leq x^\sharp}
  & \sum_{l\in[L]}\sum_{i\in [I_l]} f^{(l)}(x; p^{(l)}_i)\\\nonumber
  \st\; &\forall m\in[M]:\\\nonumber
  &\begin{aligned}[t]
    \left[g^{(m)}(x; q_j)\right]_{j\in [J_m]} +\sum_{n\in [N_m]}\sum_{k\in [K_n]}h^{(n)}(x; s^{(n)}_{k}) =0,
  \end{aligned}
  \end{align}
where $f^{(\ell)}(\cdot,\cdot)$, $g^{(m)}(\cdot,\cdot)$, and
$h^{(n)}(\cdot,\cdot)$ are twice differentiable functions with respect
to the first argument, whereas $\{\{p^{(k)}_i\}_{i\in [N_k]}\}_{k\in[K]}$,
$\{\{q^{(k)}_{i}\}_{i\in [M_l]}\}_{m\in[M]}$, and
$\{\{\{s^{(n)}_{k}\}_{k\in[K_n]}\}_{n\in[N_m]}\}_{m\in[M]}$ are
problem data, which can either be discrete or continuous.  We assume
that our functions $f^{(l)}(\cdot,\cdot)$, $g^{(m)}(\cdot,\cdot)$, and
$h^{(n)}(\cdot,\cdot)$ can be expressed with computational
graphs of moderate length. One can observe that the problem in
\eqref{eqn:prob} is expressed by the computational patterns in Section
\ref{sec:gpu}. In particular, the objective function falls within \ref{eqn:pattern-2},
the first term in the constraint falls within \ref{eqn:pattern-1}, and the second term in the constraint falls within \ref{eqn:pattern-3}.
Accordingly, the evaluation and differentiation of the
model equations in \eqref{eqn:prob} is amenable to SIMD parallelism.

@@

\subsection{Parallel AD}
\label{sec:simd:ad}
Many physics-based models, such as AC OPF, have a highly repetitive
structure. One of the manifestations of it is that the mathematical
statement of the model is concise, even if the practical model may contain
millions of variables and constraints. This is possible due to the use of
repetition over a certain index and data sets. For example,
it suffices to use 15 computational patterns to fully specify the
AC OPF model. These patterns arise from (1) generation cost, (2) reference
bus voltage angle constraint, (3-6) active and reactive power flow (from and to),
(7) voltage angle difference constraint, (8-9) apparent
power flow limits (from and to), (10-11) power balance equations,
(12-13) generators' contributions to the power balance equations, and
(14-15) in/out flows contributions to the power balance
equations. However, such repetitive structure is not well exploited in
the standard NLP modeling paradigms. In fact, without the SIMD
abstraction, it is difficult for the AD package to detect the
parallelizable structure within the model, as it will require the full
inspection of the computational graph over all expressions.  By
preserving the repetitive structures in the model, the repetitive
structure can be directly available in AD implementation.
